---
title: "Identify all possible PFAMs in your genome and extract usefull information"
output: github_document
---

##### Looking for the PFAM domains.

- Install [MEBs](https://github.com/valdeanda/mebs).

- Get into your **MEBs** directory

- Within the MEBs directory, create a directory call **pfam** within the cycles dir. 

- Move the following files to your directory cycles/pfam: [**entropies.tab**](https://github.com/mirnavazquez/Find_all_pfams/blob/master/data/), **my_Pfam.pfam.hmm** and [**pfam2kegg.tab**](https://github.com/mirnavazquez/Find_all_pfams/blob/master/data/). 

- Move into the config directory and add the next line into the end of the **config.txt**.

```{bash Edit config, eval=FALSE}
pfam	cycles/pfam/	cycles/pfam/pfam2kegg.tab	1	1	1	1	1	1	1	1
```

- Run MEBs.

```{bash Runing MEBs, eval=FALSE}
perl mebs.pl -input /path/to/the/genomes/ -type genomic -comp > out_file.tsv
```

- Explore the output file. 
- Make sure that all the PFAMs where calculated correctly for all the genomes. To do that we can check that all the genomes got an **ok** message. In the case a genome is missing you may want to run again MEBs.

```{bash Checking output file, eval=FALSE}
tail /path/to/the/genomes/*pfam.hmmsearch.tab | grep "ok" | wc -l
```

##### Make the a clustering analysis

```{bash Clustering with MEBs, eval=FALSE}
python mebs/scripts/groups_1_mod.py out_file.tsv
```

##### Which PFAMs are present in a certain cluster but not in the rest? 

We read the *.tsv file within R. As well, we load a file that contains your genome ID and to which group/cluster it corresponds. 

```{r, eval=FALSE}
library(vroom)
pfam_matrix<-vroom("out_pfam_2.tsv", delim="\t")
```

```{r include=FALSE}
load("/media/jorgers/Seagate Backup Plus Drive/02.CP9/GitProjects/Find_all_pfams/data/Cluser.RData")
```

```{r, eval=FALSE}
clusters<-vroom("PFAM_clusters", delim="\t")
```

```{r}
head(clusters)
```

```{r, eval=FALSE}
pfam_matrix_clusters<-dplyr::right_join(pfam_matrix, clusters, by = "Genomes")
```

Now run the function **get_unique_pfams**. This function that takes the PFAM data frame and parse the information based on any user group. It sums all the values of each column for a given group and the rest. If the group's sum is greater than 0, and if the sum of the rest is equal to 0, we assume that the PFAM in that column is unique for that group. 

```{r, eval=FALSE}
get_unique_pfams<-function(df_pfam, groupOfInterest){
  gruop_1<-dplyr::filter(df_pfam, Groups == groupOfInterest)
  allGroupNames <- unique(df_pfam$Groups)
  otherGroups <- allGroupNames[which(allGroupNames  != groupOfInterest )]
  The_rest <- dplyr::filter(df_pfam, Groups %in%  otherGroups)
  uniquePFAMs <- c()
  for(i in 2:(length(gruop_1)-2)){
    if(isTRUE(sum(gruop_1[,i]) > 0 & sum(The_rest[,i]) == 0)){
      uniquePFAMs <- c(uniquePFAMs, colnames(gruop_1[i]))
    }
  }
  return(uniquePFAMs)
}

```

We use that function to extract the unique PFAMs for each group. 

```{r, eval=FALSE}
Pfam_group1<-get_unique_pfams(pfam_matrix_clusters, "Group 1")
Pfam_group2<-get_unique_pfams(pfam_matrix_clusters, "Group 2")
```

##### What does this PFAM groups mean?

If you look to your unique PFAMs you will see something like this:

```{r}
head(Pfam_group1)
```

Now we need to map those IDs to meaningful names. To do that, we will use the [**pfam_mapping.txt**](https://github.com/mirnavazquez/Find_all_pfams/blob/master/data/).

```{r, eval=FALSE}
pfam_mapping<-vroom::vroom("pfam_mapping.txt")
Project_pfams<-as.data.frame(colnames(pfam_matrix))
New_Project_pfams<-as.data.frame(Project_pfams[-1,])
colnames(New_CP9_pfams) <- c("pfam_PATHWAY")
New_Project_pfams$pfam_PATHWAY<-as.character(New_Project_pfams$pfam_PATHWAY)
Mapped_pfams<-dplyr::left_join(New_Project_pfams, pfam_mapping, by = "pfam_PATHWAY")
write.table(Mapped_pfams, file="Mapped_pfams.txt", quote = F, col.names = T, row.names = F, sep = "\t")
```

```{r include=FALSE}
Mapped_pfams<-Mapped_CP9_pfams
```

```{r}
head(Mapped_pfams)
```

To extract the real name of the PFAM, we can run the script  [**pfam.terms.sh**](https://github.com/mirnavazquez/Find_all_pfams/blob/master/bash/)

##### Now the enrichement

To make an enrichment analysis, you can use the [**dcGOR**](https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1003929) package.

```{r, eval=FALSE}
library(dcGOR)
library(tibble)
```

We manipulate the files that we previously generate to create the input files to make the enrichment analysis with this function. 

```{r, eval=FALSE}
get_input_for_enrch<-function(group_list, mapping_file){
  group_list_mapped<-as_tibble(group_list)
  colnames(group_list_mapped) <- c("pfam_PATHWAY")
  Mapped_pfams<-dplyr::left_join(group_list_mapped, mapping_file, by = "pfam_PATHWAY")
  return(Mapped_pfams)
  }
```

```{r, eval=FALSE}
Enriched1<-get_input_for_enrch(Pfam_group1, pfam_mapping)
Enriched2<-get_input_for_enrch(Pfam_group2, pfam_mapping)
```

```{r, eval=FALSE}
Pfam <- dcRDataLoader('Pfam')
```

```{r, eval=FALSE}
eoutput1 <-dcEnrichment(Enriched1$PFAM, background = Mapped_CP9_pfams$PFAM, domain = 
"Pfam", ontology="GOMF")
eoutput2 <-dcEnrichment(Enriched2$PFAM, background = Mapped_CP9_pfams$PFAM, domain = 
"Pfam", ontology="GOMF")
```
